Artificial Ecosystem Selection: Methane Oxidation

### 2020-07-16 Manuscript/manuscript.Rmd
Add deviance figure and model results

---
title: "Artificial ecosystem selection on soil methane oxidation reveals simple/complex mapping between microbial community structure and ecosystem function"
author: "Andrew H. Morris and Brendan J. M. Bohannan"
output:
  html_document: default
  word_document:
    reference_docx: reference.docx
---

```{r options, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
library(tidyverse)
library(captioner)
library(broom)
library(broman)
figures <- captioner()
figures("selection", "Rank CH~4~ oxidation rate for each jar in each passage.")
figures("herit", "Correlation between mid-parent CH~4~ flux and mid-offspring CH~4~ flux.")

@@

@@

```{r sel_fig, fig.width=5, fig.height=4, fig.cap=figures("selection")}
fluxes %>% 
  ggplot(mapping = aes(x = passage, y = estimate_rank, color = treat)) +
  geom_jitter(width = 0.2) +
  geom_smooth(method = 'lm') +
  labs(x = "Passage Number", y = "Rank Methane Oxidation Rate (-k)") +
  scale_color_discrete(name = "Selection\nTreatment", labels = c('Neutral', 'Positive')) +
  theme_bw()
```

```{r herit_model}
heritability <- read_tsv('../Output/heritability.tsv')
fit <- lm(offspring ~ parental, data = heritability)
her_sum <- tidy(summary(fit))
her_aov <- tidy(anova(fit))
```

Methane oxidation rate was heritable between consecutive generations (`r figures("herit", display="cite")`). Specifically, the narrow-sense heritability ($h^2$), calculated as the slope of the mid-parent and mid-offspring, was `r myround(her_sum[2, 2], 2)`.

```{r herit_fig, fig.width=5, fig.height=4, fig.cap=figures("herit")}
ggplot(heritability, aes(x = parental, y = offspring)) +
  geom_point(aes(color = treat)) +
  stat_smooth(method = 'lm') + 
  labs(x = "Parental Mean Flux (-k)", y = "Offspring Mean Flux (-k)") +
  scale_color_discrete(name = "Selection\nTreatment", labels = c('Neutral', 'Positive')) +
  theme_bw()
```

```{r inbreeding_coef}
N = 12 # population size
T = 5  # number of passages (generations)
F = 1 - (1 - 1 / (2 * N)) ^ T
F
```

>>>>

---
title: Artificial ecosystem selection on soil methane oxidation reveals simple/complex
  mapping between microbial community structure and ecosystem function
author: "Andrew H. Morris and Brendan J. M. Bohannan"
output:
  html_document: default
  word_document:
    reference_docx: reference.docx
editor_options:
  chunk_output_type: inline
---

```{r options, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
library(tidyverse)
library(captioner)
library(broom)
library(broman)
library(morris)
library(metafor)
figures <- captioner()
figures("selection", "Rank CH~4~ oxidation rate for each jar in each passage.")
figures("deviance", "Deviance of CH~4~ oxidation rate for the Positive selection treatment relative to the Neutral selection treatment. Deviance is calculated as the mean of P within each passage minus the mean of N within each passage. Errorbars are the sums of the standard errors for P and N for each passage. The best fit line is an ordinary least squares fit taking into account the standard errors of the deviations. The gray ribbon is the 95% confidence interval for the regression coefficient.")
figures("herit", "Correlation between mid-parent CH~4~ flux and mid-offspring CH~4~ flux. The mid-parents are the mean of the three (or two) jars selected to inoculate the next passage. The mid-offsprings are the mean of all twelve jars produced in one passage. The best fit line is an ordinary least squares regression line and the gray ribbon is the 95% confidence interval for the regression coefficient.")

@@

## Response to selection

@@

```{r sel_fig, fig.width=5, fig.height=3, fig.cap=figures("selection")}
fluxes %>% 
  ggplot(mapping = aes(x = passage, y = estimate_rank, color = treat)) +
  geom_jitter(width = 0.2) +
  geom_smooth(method = 'lm', formula = y ~ x, se = FALSE) +
  labs(x = "Passage Number", y = "Rank Methane Oxidation Rate (-k)") +
  #scale_color_discrete(name = "Selection\nTreatment", labels = c('Neutral', 'Positive')) +
    scale_color_manual(name = "Selection\nTreatment", labels = c('Neutral', 'Positive'), values = c("gray60", "darkorange2")) +
  theme_bw()
```

## Deviance

```{r dev_model}
deviance_se <- 
  fluxes %>% 
  select(treat, passage, estimate) %>% 
  group_by(passage, treat) %>% 
  summarize(se = se(estimate)) %>% 
  pivot_wider(names_from = treat, values_from = se) %>% 
  mutate(deviance_se = p + n) %>% 
  pull(deviance_se)

deviance <-
fluxes %>% 
  select(treat, passage, estimate) %>% 
  group_by(passage, treat) %>% 
  summarize(mean = mean(estimate)) %>% 
  pivot_wider(names_from = treat, values_from = mean) %>% 
  mutate(deviance = p - n)

deviance$se <- deviance_se

dev_fit <- rma(deviance~passage, se^2, method="FE", data = deviance)
dev_fit1 <- lm(deviance~passage, data = deviance)

flux_per_psg <-
fluxes %>% 
  group_by(passage) %>% 
  summarize(mean = mean(estimate), se = se(estimate)) %>% 
  pull(mean)

fluxes %>% 
  group_by(passage, treat) %>% 
  summarize(mean = mean(estimate), se = se(estimate))

dev_pred <- predict(dev_fit)
```

Methane oxidation rate increased as a response to selection in the Positive selection treatment (`r figures("deviance", display="cite")`). This can be seen in the increase in deviation between the Positive and Neutral selection treatments over the course of the experiment (z = `r myround(dev_fit$zval[2], 2)`, p = `r myround(dev_fit$pval[2], 3)`). Mean CH~4~ oxidation rate for the Positive treatment increased by `r myround(dev_fit$beta[2], 3)` per passage relative to the Neutral treatment. The mean flux across all jars in all treatments in the first passage was `r myround(flux_per_psg[1], 3)` so this effectively represents a doubling of CH~4~ consumption in every passage due to selection.

```{r dev_fig, fig.width=4, fig.height=3, fig.cap=figures("deviance")}

ggplot(deviance, aes(x = passage, y = deviance, ymin = deviance - se, ymax = deviance + se)) + 
  geom_ribbon(aes(ymin = dev_pred$ci.lb, ymax = dev_pred$ci.ub), alpha = 0.2) +
  geom_pointrange(color = 'darkorange2') + 
  #geom_abline(intercept = 0.1 + fit$beta[1], slope = fit$beta[2], color = 'blue') +
      geom_segment(aes(x = 1, xend = 5, y = dev_fit$beta[1] + dev_fit$beta[2], yend = dev_fit$beta[1] + dev_fit$beta[2]*5), color = 'darkorange2', size = 1) +
  labs(x = "Passage Number", y = "Deviation from Control (P - N)")+
  theme_bw()
```

## Ecosystem heritability

```{r herit_model}
heritability <- read_tsv('../Output/heritability.tsv')
her_fit <- lm(offspring ~ parental, data = heritability)
her_sum <- tidy(summary(her_fit))
her_aov <- tidy(anova(her_fit))

```{r inbreeding_coef}
N = 12 # population size
T = 5  # number of passages (generations)
F = 1 - (1 - 1 / (2 * N)) ^ T
```

Methane oxidation rate was heritable between consecutive passages (`r figures("herit", display="cite")`). Specifically, the narrow-sense heritability, calculated as the slope of the mid-parent and mid-offspring, was $h^2$ = `r myround(her_sum[2, 2], 2)` $\pm$ `r myround(her_sum[2, 3], 2)`. With an inbreeding coefficient of `r myround(F, 2)`,

```{r herit_fig, fig.width=5, fig.height=4, fig.cap=figures("herit")}
ggplot(heritability, aes(x = parental, y = offspring)) +
  geom_point(aes(color = treat)) +
  stat_smooth(method = 'lm', formula = y ~ x, color = "black") + 
  labs(x = "Parental Mean Flux (-k)", y = "Offspring Mean Flux (-k)") +
    scale_color_manual(name = "Selection\nTreatment", labels = c('Neutral', 'Positive'), values = c("gray60", "darkorange2")) +
  theme_bw()

```{r}
R = dev_fit$beta[2]
h2 = pull(her_sum[2, 2])
S = R/h2
```

```{r}
fluxes
heritability
```

### 2020-07-16 R/process.Rmd
Add deviance figure and model results

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,  fig.align="center")
library(tidyverse)
library(knitr)
library(kableExtra)
library(broom)
library(modelr)
library(lubridate)
library(morris)
library(magrittr)
```

@@

sc %>% 
  filter(molecule == 'ch4') %>% 
ggplot(mapping = aes(area, injection_ppm)) +
  facet_wrap(~ flux_date + molecule, scales = "free") + 
  geom_point() +
  stat_smooth(method = 'lm', se = FALSE)

@@

fit <- lm(estimate ~ treat * passage, data = fluxes)
summary(fit)
plot(fit$residuals ~ fluxes$estimate)
selection_table <- tidy(anova(fit))

@@

fluxes %>% 
  ggplot(mapping = aes(x = passage, y = estimate, color = treat)) +
  geom_jitter(width = 0.2) +
  #geom_smooth(method = 'lm') +
  labs(x = "Passage Number", y = "Methane Oxidation Rate (-k)") +
  scale_color_discrete(name = "Selection Treatment", labels = c('Neutral', 'Positive')) +
  theme_bw() + 
  geom_boxplot(aes(x = passage, y = estimate, color = treat, group = interaction(treat, passage)))

@@

deviance <-
fluxes %>% 
  select(treat, passage, estimate) %>% 
  group_by(passage, treat) %>% 
  summarize(mean = median(estimate)) %>% 
  pivot_wider(names_from = treat, values_from = mean) %>% 
  mutate(deviance = p - n)

@@

ggplot(deviance, aes(x = passage, y = deviance, ymin = deviance - se, ymax = deviance + se)) + 
  geom_pointrange() + 
  geom_abline(intercept = fit$beta[1], slope = fit$beta[2], color = 'blue')

@@

```{r}
library(rstan)

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,  fig.align="center")
library(tidyverse)
library(knitr)
library(kableExtra)
library(broom)
library(modelr)
library(lubridate)
library(morris)
library(magrittr)
library(metafor)
```

@@

sc %>% 
  filter(molecule == 'ch4') %>% 
ggplot(mapping = aes(area, injection_ppm)) +
  facet_wrap(~ flux_date + molecule, scales = "free") + 
  geom_point() +
  stat_smooth(formula = y ~ x, method = 'lm', se = FALSE)

@@

fit <- lm(estimate ~ treat * passage, data = fluxes)
plot(fit$residuals ~ fluxes$estimate)
selection_table <- tidy(anova(fit))

@@

fluxes %>% 
  ggplot(mapping = aes(x = passage, y = estimate, color = treat)) + 
  geom_boxplot(aes(x = passage, y = estimate, color = treat, group = interaction(treat, passage))) +
  geom_jitter(width = 0.2) +
  #geom_smooth(method = 'lm') +
  labs(x = "Passage Number", y = "Methane Oxidation Rate (-k)") +
  scale_color_discrete(name = "Selection Treatment", labels = c('Neutral', 'Positive')) +
  theme_bw()

@@

deviance <-
fluxes %>% 
  select(treat, passage, estimate) %>% 
  group_by(passage, treat) %>% 
  summarize(mean = mean(estimate)) %>% 
  pivot_wider(names_from = treat, values_from = mean) %>% 
  mutate(deviance = p - n)

@@

ggplot(deviance, aes(x = passage, y = deviance, ymin = deviance - se, ymax = deviance + se)) + 
  geom_pointrange(color = 'darkorange2') + 
  geom_abline(intercept = fit$beta[1], slope = fit$beta[2], color = 'blue')

@@

```{r eval=FALSE, include=FALSE}
library(rstan)
